Standardization of ceramic shape: A case studay from the Iron Age pottery from northeastern Taiwan

### 2020-07-24 analysis/paper/paper.Rmd
update description file

```{r size-regression, fig.cap= "A: Distribution of body diamater of vessels across three phases. B: Scatter plots of shape variable, pc1 at the top and pc2 at the bottom, as a function of body dimater"}
library(ggpubr)
library(geomorph)

@@

Previous investigations at Kiwulan suggested an unequal distribution of prestige goods, trade ornaments in particular, following the appearance of Europeans [@Cheng2008; @Wang2011], which hinted at the emergence of social inequality within the Indigenous community. Ceramic vessel shape standardization was examined to measure craft specialization as a proxy for social differentiation [@Costin2001; @Junker1999]. The result of our MANOVA demonstrates a significant difference in shapes between the pre-European and Post-European periods, and between the pre-European and the Chinese periods. The average shape presents as a round body with a wide rim and neck before European contact, which shifts to a more oval-shaped body with narrower rim and neck after the European presence. Shape changes are more pronounced during the Chinese contact period. In general, vessels become shorter in height over time, leading to an oval-shaped body. Moreover, size significantly varies with shape for vessels before and after the European presence, which presents that shorter vessels with narrower opening tend to have a larger body diameter. This correlation between narrower opening and larger diameter can be also observed during the Chinese period without the effect of height.

>>>>

```{r size-regression, fig.cap= "A: Distribution of the body diameter of the vessels across three phases. B: Scatter plots of shape variable, PC1 at the top and PC2 at the bottom, as a function of body diamater"}

library(ggpubr)
library(geomorph)

@@

Previous investigations at Kiwulan suggested an unequal distribution of prestige goods, trade ornaments in particular, following the appearance of Europeans [@Cheng2008; @Wang2011], which hinted at the emergence of social inequality within the Indigenous community. Ceramic vessel shape standardization was examined to measure craft specialization as a proxy for social differentiation [@Costin2001; @Junker1999]. The result of our MANOVA demonstrates a significant difference in shapes between the pre-European and Post-European periods, and between the pre-European and the Chinese periods. The average shape presents as a round body with a wide rim and neck before European contact, which shifts to a more oval-shaped body with narrower rim and neck after the European presence. Shape changes are more pronounced during the Chinese contact period. In general, vessels become shorter in height over time, leading to an oval-shaped body. Moreover, size significantly varies with shape for vessels before and after the European presence, which presents that shorter vessels with narrower opening tend to have a larger body diameter. This correlation between size and shape can be also observed during the Chinese period without the consideration of height.
